\Large \bf Vigor: Toward Carrier-Grade Software Network Functions

### 2016-11-22 doc/paper/paper.tex
port != 9 condition jorney accross the Vigor framework.

The ring keeps the state of our NF and makes many existing verification
approaches impractical.

@@

\begin{figure}[h!]
\begin{lstlisting}
int main(int argc, char** argv) {
  struct packet p;
  struct ring *r = ring_create(CAP);^\label{line:mk-ring}^
  if (!r) return 1;
  while(LOOP(1))^\label{line:loop-hd}^
  {
    loop_iteration_begin(&r);
    if (!ring_full(r))
      if (receive_packet(&p) && p.port != 9)
        ring_push_back(r, &p);
    if (!ring_empty(r) && can_send_packet()) {
      ring_pop_front(r, &p);
      VIGOR_CHECK(p.port != 9);^\label{line:vigcheck}^
      send_packet(&p);
    }
    loop_iteration_end(&r);
  }
  return 0;
}
\end{lstlisting}
  \caption{Vigor implementation of Drop protocol.
    \code{LOOP}(l.\ref{line:loop-hd}) labels the infinite loop head and we
    explicitly use a ring(l.\ref{line:mk-ring}) from the Vigor library. \code{VIGOR_CHECK} at the
    line \ref{line:vigcheck} adds a semantic property to be verified.}
  \label{lst:vigor}
\end{figure}

@@

/*@^\label{line:vf-user-prop-from}^
  fixpoint bool
  user_property(packet p) {
    switch(p) { case packet(port):
                  return port != 9; }
  }
  @*/^\label{line:vf-user-prop-to}^

static bool^\label{line:c-user-prop-from}^
packet_constraints(struct packet* p) {
  return p->port != 9;
}^\label{line:c-user-prop-to}^
\end{lstlisting}
  \caption{Custom definitions capturing the desired property. They capture the
    essence of every packet in the ring. Currently, we have to duplicate the
    packet condition in two languages (formal logic --
    ll\ref{line:vf-user-prop-from}-\ref{line:vf-user-prop-to}, and C --
    ll\ref{line:c-user-prop-from}-\ref{line:c-user-prop-to}) due to the
    limitations of our prototype.}
  \label{lst:customizations}
\end{figure}

@@

This change makes SE feasible, albeit impractical: for \code{CAP=5}, it takes 10
sec to exhaustively SE the code; for \code{CAP=30} it takes 15 min, for
\code{CAP=60} it takes 3 hours, and so on. Another classic SE challenge ---
``symbolic
indexing''~\cite{sen2005cute,godefroid2008automated,boonstoppel2008rwset} spoils
the party here. When
indexing an array with a symbolic variable (or when dereferencing a symbolic
pointer), the constraints on program state become extremely complex, because the
symbolic pointer or index could be referencing many different possible
locations. The complexity of the constraints chokes the SMT solver, and the SEE
grinds to a halt. In general, SE doesn't work for code with symbolic pointers or
offsets. Alas, stateful C code almost always has some of these.

In this example the symbolic indexing comes from the ring array. Once we
replaced the infinite loop with a single iteration we have to assume (remember
havoc-ing) the \code{begin} and \code{begin + queue_len} indexes may point
anywhere in the array.

@@

\begin{figure}[h!]
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p) {
  FILL_SYMBOLIC(p, sizeof(struct packet),
                "popped_packet");
  ASSUME(packet_constraints(p));
}
\end{lstlisting}
  \caption{A symbolic model for \code{ring_pop_front}. It fills symbolically the
    a packet and adds a constraint to it.}
  \label{lst:symbolic-model}
\end{figure}

@@

A model of \code{ring_pop_front} used for this example is shown on
Fig.~\ref{lst:symbolic-model}. This model is simple enough to enable
exhaustive SE, but preserves the vital property --- \code{packet_constraints}
(Fig.\ref{lst:customizations}).

@@

/*@ assert(arg2.port != 9); @*/
\end{lstlisting}
  \caption{A call sequence, produced by SEE. Vigor adds formal annotations to
    constrain the execution space.}
  \label{lst:call-sequence}
\end{figure}

@@

>>>>

Our NF accumulates state in the ring thus making many existing verification
approaches impractical.

@@

\begin{figure}[h!]
\begin{lstlisting}
int main(int argc, char** argv) {
  struct packet p;
  struct ring *r = ring_create(CAP);^\label{line:mk-ring}^
  if (!r) return 1;
  while(LOOP(1))^\label{line:loop-hd}^
  {
    loop_iteration_begin(&r);
    if (!ring_full(r))
      if (receive_packet(&p) && p.port != 9)^\label{line:if-cond}^
        ring_push_back(r, &p);
    if (!ring_empty(r) && can_send_packet()) {
      ring_pop_front(r, &p);^\label{line:pop-front}^
      VIGOR_CHECK(p.port != 9);^\label{line:vigcheck}^
      send_packet(&p);
    }
    loop_iteration_end(&r);
  }
  return 0;
}
\end{lstlisting}
  \caption{Vigor implementation of Drop protocol.
    \code{LOOP}(l.\ref{line:loop-hd}) labels the infinite loop head and we
    explicitly use a ring(l.\ref{line:mk-ring}) from the Vigor library. \code{VIGOR_CHECK} at the
    line \ref{line:vigcheck} adds a semantic property to be verified.}
  \label{lst:vigor}
\end{figure}

@@

/*@^\label{line:vf-user-prop-from}^
  fixpoint bool
  user_property(packet p) {
    switch(p) { case packet(port):
                  return port != 9; }^\label{line:vf-user-prop}^
  }
  @*/^\label{line:vf-user-prop-to}^

static bool^\label{line:c-user-prop-from}^
packet_constraints(struct packet* p) {
  return p->port != 9;^\label{line:c-pkt-cnstr}^
}^\label{line:c-user-prop-to}^
\end{lstlisting}
  \caption{Custom definitions capturing the desired property. They capture the
    essence of every packet in the ring. Currently, we have to duplicate the
    packet condition in two languages (formal logic --
    ll\ref{line:vf-user-prop-from}-\ref{line:vf-user-prop-to}, and C --
    ll\ref{line:c-user-prop-from}-\ref{line:c-user-prop-to}) due to the
    limitations of our prototype.}
  \label{lst:customizations}
\end{figure}

@@

This change makes SE feasible, albeit impractical: for \code{CAP=5}, it takes 10
sec to exhaustively SE the code; for \code{CAP=30} it takes 15 min, for
\code{CAP=60} it takes 3 hours, and so on. Another classic SE challenge ---
``symbolic
indexing''~\cite{sen2005cute,godefroid2008automated,boonstoppel2008rwset} spoils
the party here. When indexing an array with a symbolic variable (or when
dereferencing a symbolic pointer), the constraints on program state become
extremely complex, because the symbolic pointer or index could be referencing
many different possible locations. The complexity of the constraints chokes the
SMT solver, and the SEE grinds to a halt. In general, SE is not practical for
code with symbolic pointers or offsets. Alas, stateful C code almost always has
some of these.

In this example the symbolic indexing comes from the ring array. Once we
replaced the infinite loop with a single iteration we have to assume (due to
havoc-ing) the \code{begin} and \code{begin + queue_len} indexes may point
anywhere in the array.

@@

\begin{figure}[h!]
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p) {
  FILL_SYMBOLIC(p, sizeof(struct packet),
                "popped_packet");
  ASSUME(packet_constraints(p));^\label{line:assume}^
}
\end{lstlisting}
  \caption{A symbolic model for \code{ring_pop_front}. It fills symbolically the
    a packet and adds a constraint to it.}
  \label{lst:symbolic-model}
\end{figure}

@@

The model of \code{ring_pop_front} used for this example is shown on
Fig.~\ref{lst:symbolic-model}. This model is simple enough to enable
exhaustive SE, but preserves the vital property --- \code{packet_constraints}
(Fig.\ref{lst:customizations}).

@@

/*@ assert(arg2.port != 9); @*/^\label{line:assert}^
\end{lstlisting}
  \caption{A call sequence, produced by SEE. Vigor adds formal annotations to
    constrain the execution space.}
  \label{lst:call-sequence}
\end{figure}

@@

Let us follow now the `pord != 9' condition across the framework. During
exhaustive SE, SEE first learns the constraint in the \code{if} condition:
Fig.~\ref{lst:vigor} l.\ref{line:if-cond}. It however forgets it on the next
line, because our model of \code{ring_push_back} discards the packet being
pushed. However, this constraint arises again on the call to
\code{ring_pop_front}(l.\ref{line:pop-front}) as
\code{packet_constraints}(Fig.~\ref{lst:customizations},l.\ref{line:c-pkt-cnstr}),
according to the symbolic model - Fig.~\ref{lst:symbolic-model}. Being applied
to the fresh packet from the ring, it satisfies the \code{VIGOR_CHECK}
(Fig.~\ref{lst:vigor} l.\ref{line:vigcheck}).

During the operation, SEE produces call sequences that also capture the said
constraint. E.g. the line~\ref{line:assert} on Fig.~\ref{lst:call-sequence}
reflects the condition (Fig.~\ref{lst:customizations} l.\ref{line:c-pkt-cnstr})
that was added by the symbolic model (Fig.~\ref{lst:symbolic-model}
l.\ref{line:assume}).

Formal verification serves two purposes:
\begin{itemize}
\item verification of the ring implementation. During this part, it proves that
  \code{ring_pop_front} guarantees the popped packet to satisfy the
  \code{user_property} (Fig.~\ref{lst:contract} l.\ref{line:postcondition-to}).
\item validation of the call sequence. Here it uses the specific user property -
  Fig.~\ref{lst:customizations} l.\ref{line:vf-user-prop}, to check it against
  the assertion of the \code{ring_pop_front} output:
  Fig.~\ref{lst:call-sequence} l.\ref{line:assert}.
\end{itemize}
